---
title: "Adaptive Rejection Metropolis Sampling in R"
author: "Michael Bertolacci"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adaptive Rejection Metropolis Sampling in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(armspp)
```

# Background

# Gibbs sampling

# Using the R package

## Overview

## Example: normal distribution

A very simple example, for which exact sampling is possible, is the unit normal
distribution. This has the unnormalised log density of $-\frac{x^2}{2}$, with
the entire real line as its domain, and the density is log concave. This means
we can use `metropolis = FALSE` to get exact independent samples:

```{r normal_example}
output <- arms(
  5000, function(x) -x ^ 2 / 2,
  -1000, 1000,
  metropolis = FALSE, include_n_evaluations = TRUE
)
print(str(output))
hist(output$samples, br = 50, freq = FALSE, main = 'Normal samples')
shapiro.test(output$samples)
```

(Note: there are obviously better ways to sample this distribution---`rnorm`
for a start.)

## Example: mixture of normal distributions

Another simple example, but one for which the Metropolis-Hastings step is
required, is a mixture of normal distributions. For instance, consider

$$
  x \sim 0.4 N(-1, 1) + 0.6 N(4, 1),
$$

which has a density that looks like

```{r mixture_of_normals_density}
dnormmixture <- function(x) {
  parts <- log(c(0.4, 0.6)) + dnorm(x, mean = c(-1, 4), log = TRUE)
  log(sum(exp(parts - max(parts)))) + max(parts)
}

curve(exp(Vectorize(dnormmixture)(x)), -7, 10)
```

This distribution is not log-concave, so we need to use `metropolis = TRUE` to
correct the inexactness caused by the use of an imperfect rejection
distribution. Doing this we can sample from the distribution

```{r mixture_of_normals_example}
samples <- arms(1000, dnormmixture, -1000, 1000)
hist(samples, freq = FALSE, br = 50)
curve(exp(Vectorize(dnormmixture)(x)), -7, 10, col = 'red', add = TRUE)
```

# Using the C++ implementation